\textbf{Sharp Bias-variance Tradeoffs of Hard Parameter Sharing in High-dimensional Linear Regression}

### 2020-12-03 AISTATS2021_Author_Response_Pack/aistats2021_author_response.tex

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % define colors in text
\usepackage{xspace}         % fix spacing around commands

@@

	\textbf{Response to R1.} Please see our response to the detailed comments.
	\textbf{(8-i)} Yes, the variance decreases because $||{a_i^{\star}}||^2 \le 1$.
	The bias increases because the bias of single-task learning (the ordinary least squares solution) is zero when $n > p$.
	\textbf{(8-ii)} By lower order terms, we mean any term that vanishes to zero when $p$ increases to infinity.
	\textbf{(8-iii)} The sum of $||{a_i^{\star}}||^2$ is one (or $r$ in general), because $A^{\star}$ is a (subspace) projection matrix.
	\textbf{(8-iv)} %We want to point out that the problem setting in Section 3 is in general non-convex. Therefore, it is not possible to explicitly solve $A_1, A_2$ without making further assumptions.
	We can assume knowledge of $A_1 \in \mathbb R, A_2 \in \mathbb R$ without loss of generality since they can be solved numerically by using a subset of training data points (e.g. $\sqrt{p}$ suffices).
	The sampling error of the numerical solution is $O(\exp(-p))$.

	\textbf{Response to R2.} We thank the reviewer for providing detailed feedback.
	Response to questions under Section 8.
	\textbf{(i)}
	In the over-parametrized regime, the minimum solution of $f(A ,B)$ (equation 1) reduces to single-task learning.
	Hence, studing this regime is problematic without adding explicit regularization or implicit algorithmic biases.
	These are interesting research questions, but beyond the scope of our work.
	\textbf{(ii)}

	Response to questions under Section 3.
	\textbf{(1-i)}

	\textbf{Response to R3.}

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % define colors in text
\usepackage{xspace}         % fix spacing around commands
\usepackage{macro_math}
\usepackage{macro_fan}

@@

	\textbf{Response to R1.} Please see our response to the detailed comments.
	\textbf{(8-i)} Yes, the variance decreases because $||{a_i^{\star}}||^2 \le 1$.
	The bias increases because the bias of single-task learning (the ordinary least squares solution) is zero when $n > p$.
	\textbf{(8-ii)} By lower order terms, we mean any term that vanishes to zero when $p$ increases to infinity.
	\textbf{(8-iii)} The sum of $||{a_i^{\star}}||^2$ is one because $A^{\star}$ is a rank-$1$ projection matrix.
	\textbf{(8-iv)} %We want to point out that the problem setting in Section 3 is in general non-convex. Therefore, it is not possible to explicitly solve $A_1, A_2$ without making further assumptions.
	We can assume knowledge of $A_1 \in \mathbb R, A_2 \in \mathbb R$ without loss of generality since they can be solved numerically by using a subset of training data points (e.g. $\sqrt{p}$ suffices).
	The sampling error of the numerical solution is $O(\exp(-p))$.

	\textbf{Response to R2.} We thank the reviewer for providing detailed feedback.
	Response to questions under Section 8.
	\textbf{(i)}
	In the over-parametrized regime, the minimum solution of $f(A ,B)$ (equation 1) reduces to single-task learning  (cf. Proposition 1 in Wu et al., 20).
	Hence, studing this regime is problematic without adding explicit regularization or implicit algorithmic biases.
	These are interesting research questions, but beyond the scope of our work.

	Response to questions under Section 3.
	\textbf{(1-ii)} We would like to reminder the reviewer that the purpose of the examples is to explain the empirical phenomena of multi-task and transfer learning.
	We believe that our examples have provided important initial insights toward understanding the effect of varying sample size ratios and covariate shifts.
	See Appendix A for evaluation in a real world data setting.
	We will address these issues and the other clarification questions in the next version of the paper.

	\textbf{Response to R3.}
	We thank the reviewer for the encouraging review.
	Regarding the question of how the prediction loss varies with $r$, we would like to point out that in the setting of Theorem 2.1, our result implies the following simple corollary.
	Suppose that the singular values of ${B^{\star}}^{\top}\Sigma B^{\star}$ are $\lambda_1, \lambda_2, \dots, \lambda_t$ in decreasing order.
	If we increase $r$ by $1$, the bias reduces by $\lambda_{r+1}$ while the variance increases, the amount of which depends on the singular vector of  ${B^{\star}}^{\top}\Sigma B^{\star}$ corresponding to $\lambda_{r+1}$.
	As an example of this corollary, we can show that in the setting of Example 2.2, the average prediction loss of all tasks is minimized when $r$ is one.
	We also thank the reviewer for suggesting the related work.
	Our work is heavily inspired by Hastie et al., 2019.
	We will add these related work to the next version of the paper.

### 2020-12-03 arxiv/conclusions.tex

\vspace{-0.06in}
\section{Conclusions and Discussions}\label{sec_conclude}
\vspace{-0.06in}

This work studied generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS, and rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.
We describe open questions for future work.
First, it would be interesting to tighten our estimate in Corollary \ref{cor_MTL_loss}, which would extend the observation in Figure \ref{fig_size} to small $n_1$.
Second, it would be interesting to extend our result to classification problems such as logistic regression.
%First, our bound in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
%Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
%This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
%FY: $+\id$ is very important and makes the problem very hard; otherwise the problem can be solved with current RMT methods.
%The eigenvalue distribution of this matrix, which has been obtained in \citet{Fmatrix}, might be helpful towards resolving this problem.
%but its singular values will follow a different distribution since the matrix is not symmetric.
 %might require new techniques beyond the current ones in random matrix theory .%\HZ{to add}.
%Second, it would be interesting to extend our results to classification problems.
%Several recent work have made remarkable progress for logistic regression in the high-dimensional  setting, e.g. \citet{sur2019modern}.
%An interesting question is to study logistic regression in a multiple-sample setting.

>>>>

\section{Conclusions and Discussions}\label{sec_conclude}

This work studied the generalization properties of a widely used hard parameter sharing approach for multi-task learning.
We provided sharp bias-variance tradeoffs of HPS in high-dimensional linear regression.
Using these results, we analyzed how varying sample sizes and covariate shifts impact HPS.
We rigorously explained several empirical phenomena such as negative transfer and covariate shift related to these dataset properties.
We validated our theory and conducted further studies on text classification tasks.

We describe several open questions for future work.
First, our result in Corollary \ref{cor_MTL_loss} involves an error term that scales down with $n_1$.
Tightening this error bound requires showing the limit of $\normFro{({Z^{(1)}}^{\top} Z^{(1)} + {Z^{(2)}}^{\top} Z^{(2)})^{-1} {Z^{(1)}}^{\top} Z^{(1)}}^2$ for two isotropic sample covariance matrices.
This requires studying the asymptotic singular values distribution of the non-symmetric matrix $({Z^{(1)}}^{\top} Z^{(1)})^{-1}{Z^{(2)}}^{\top} Z^{(2)}+\id$, which is still an open problem in random matrix theory.
The eigenvalue distribution of this matrix, which has been obtained in \citet{Fmatrix}, might help resolve this problem.
Second, it would be interesting to extend our results to classification problems.
Several recent work has made remarkable progress for logistic regression in the high-dimensional setting (e.g. \citet{sur2019modern}).
It is an interesting question to study logistic regression in a multiple-sample setting.

### 2020-12-03 arxiv/intro.tex

% The idea is to use a shared feature space for all tasks, while each task also has a separate layer for making the prediction.
\begin{abstract}
	\vspace{-0.1in}
	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties has not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
	\vspace{-0.15in}
\end{abstract}

\vspace{-0.1in}
\section{Introduction}\label{sec_introduction}

Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \citet{C97}.
Recent work has revived interest in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
Often, hard parameter sharing offers two critical advantages if successfully applied.
First, it reduces model parameters since all tasks use the same feature space.
Second, it reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

Hard parameter sharing works as an inductive transfer mechnism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12}.
Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
However, how these dataset properties impact HPS has not been well established.
%It becomes increasingly important to understand HPS' formal generalization properties.
Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
Previous results based on Rademacher complexity or VC dimensions have considered cases where all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.
%For, the generalization error scales down as the sample sizes of all tasks increase, when applied to the multi-task setting \cite{B00,AZ05,M06,MPR16,WZR20}.

\begin{figure*}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.45\textwidth,valign=t]{figures/mtl_model_arch.pdf}
		\caption{A hard parameter sharing architecture}
		\label{fig_intro_arch}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400.eps}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\vspace{-0.1in}
	\caption{
	Left: an illustrative picture of HPS.
	Right: an illustrative example of using HPS for two tasks $X_1, Y_1$ and $X_2, Y_2$ with sample size $n_1, n_2$, respectively.
	Increasing $n_1/n_2$ decreases task two's prediction loss initially, but increases afterward. This phenomenon occurs due to different bias-variance tradeoffs as $n_1/n_2$ increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
	\vspace{-0.07in}
\end{figure*}

This paper presents new techniques to study hard parameter sharing and establishes a number of new results.
We consider regression analysis, which is arguably one of the most fundamental problems in statistics and machine learning.
We are interested in the \textit{high-dimensional} setting, where each dataset's sample size and feature dimension grow linearly instead of logarithmicly.
This setting captures the fact that in many applications, a single task's sample size is usually insufficient for accurate learning.
For example, if a dataset's sample size is only a constant factor of dimension in linear regression, the variance is also constant (cf. Fact \ref{fact_tr}).
The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.

>>>>

\begin{abstract}
	Hard parameter sharing for multi-task learning is widely used in empirical research despite the fact that its generalization properties have not been well established in many cases. This paper studies its generalization properties in a fundamental setting: How does hard parameter sharing work given multiple linear regression tasks? We develop new techniques and establish a number of new results in the high-dimensional setting, where the sample size and feature dimension increase at a fixed ratio. First, we show a sharp bias-variance decomposition of hard parameter sharing, given multiple tasks with the same features. Second, we characterize the asymptotic bias-variance limit for two tasks, even when they have arbitrarily different sample size ratios and covariate shifts. We also demonstrate that these limiting estimates for the empirical loss are incredibly accurate in moderate dimensions. Finally, we explain an intriguing phenomenon where increasing one task's sample size helps another task initially by reducing variance but hurts eventually due to increasing bias. This suggests progressively adding data for optimizing hard parameter sharing, and we validate its efficiency in text classification tasks.
\end{abstract}

\section{Introduction}\label{sec_introduction}

Hard parameter sharing (HPS) for multi-task learning is widely used in empirical research and goes back to the seminal work of \citet{C97}.
Recent work has revived interests in this approach because it improves performance and reduces the cost of collecting labeled data \cite{R17}.
It is generally applied by sharing the feature layers between all tasks while keeping an output layer for every task.
Often, hard parameter sharing offers two critical advantages if successfully applied.
First, it reduces model parameters since all tasks use the same feature space.
Second, it reduces the amount of labeled data needed from each task by augmenting the entire training dataset.

Hard parameter sharing works as an inductive transfer mechanism and a regularizer that reduces overfitting, both of which have great intuitive appeal \cite{R17}.
For example, by restricting the shared space's size, HPS encourages information sharing among multiple tasks \cite{KD12,WZR20}.
Another source of inductive bias comes from the tasks and depends on datasets' properties such as sample sizes and task covariances \cite{WZR20}.
However, how these dataset properties impact HPS has not been well established.
Part of the challenge may be that HPS' generalization performance depends intricately on the sample size ratios and covariate shifts between tasks, and is not amenable to standard concentration results.
Previous results based on Rademacher complexity or VC dimensions have considered cases where all tasks' sample sizes are equal to logarithmic factors of the feature dimension \cite{B00,MPR16}, and when all tasks' sample sizes increase simultaneously \cite{AZ05,M06}.

\begin{figure*}[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.45\textwidth,valign=t]{figures/mtl_model_arch.pdf}
		\caption{A hard parameter sharing architecture}
		\label{fig_intro_arch}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\includegraphics[width=0.745\textwidth,valign=t]{figures/sample_ratio_c2_400-eps-converted-to.pdf}
		\caption{Varying sample size ratio}
		\label{fig_intro_sample_size_b}
	\end{subfigure}
	\vspace{-0.1in}
	\caption{
	Left: an illustrative picture of HPS.
	Right: an illustrative example of using HPS for two tasks $X_1, Y_1$ and $X_2, Y_2$ with sample size $n_1, n_2$, respectively.
	Increasing $n_1/n_2$ decreases task two's prediction loss initially but increase afterward. This phenomenon occurs due to different bias-variance tradeoffs as $n_1/n_2$ increases. Our result provides an estimated loss (solid line) that accurately matches the empirical loss (dotted line).
	See Section \ref{sec_simulation} for the precise setting.}
	\label{fig_intro_sample_size}
\end{figure*}

This paper presents new techniques to study hard parameter sharing and establishes a number of new results.
We consider regression analysis, which is arguably one of the most fundamental problems in statistics and machine learning.
We are interested in the \textit{high-dimensional} setting, where each dataset's sample size and feature dimension grow linearly instead of logarithmically.
This setting captures the fact that a single task's sample size is usually insufficient for accurate learning in many applications.
For example, if a dataset's sample size is only a constant factor of dimension in linear regression, the variance is also constant (cf. Fact \ref{fact_tr}).
The high-dimensional setting is challenging but is crucial for understanding how datasets' sample sizes impact generalization performance.
